---
title: "IntroductionToNeighborhoods"
author: "Zack W Almquist (University of Minnesota)"
date: "November 30, 2016"
output: html_document
---

# Preliminaries
```{r,include=FALSE}
## Useful http://www.stat.cmu.edu/~cshalizi/rmarkdown/
require(knitr)
# Set knitr options for knitting code into the report:
# - Don't print out code (echo)
# - Save results so that code blocks aren't re-run unless code changes (cache),
# _or_ a relevant earlier code block changed (autodep), but don't re-run if the
# only thing that changed was the comments (cache.comments)
# - Don't clutter R output with messages or warnings (message, warning)
  # This _will_ leave error messages showing up in the knitted report
opts_chunk$set(echo=TRUE,
               cache=TRUE, autodep=TRUE, cache.comments=FALSE,
               message=FALSE, warning=FALSE,tidy=TRUE)
options(width=60)
```

This workshop will involve a large collection of R packages and will provide a basic run through for accessing spatial data, spatial plotting and integrating with the 'ggplot' packages. Last we will perform some simple spatial statistics.

## From CRAN

### Required

```{r, eval=FALSE}
install.packages("sp")
install.packages("rgeos")
install.packages("maptools")
install.packages("spdep")
install.packages("ggplot2")
install.packages("devtools")
install.packages("plyr")
install.packages("dplyr")
install.packages("ggmap",type="source")
install.packages("png")
install.packages("rjson")
install.packages("rgeos")
```

### Recommended
```{r, eval=FALSE}
install.packages("rgdal")
```

## US Census Packages
In this workshop we will use a series of different packages, including the US Census suite of software. The most recent versions are available <https://r-forge.r-project.org/R/?group_id=2022>. We will need only three of the packages:

```{r, eval=FALSE}
install.packages("UScensus2010", repos="http://R-Forge.R-project.org")
install.packages("UScensus2010county", repos="http://R-Forge.R-project.org")
install.packages("UScensus2010tract", repos="http://R-Forge.R-project.org")
install.packages("UScensus2010cdp", repos="http://R-Forge.R-project.org")
```

## Introduction to Minneapolis

### Google Maps
```{r, cache=FALSE,fig.width = 4.5, fig.height = 4.5,fig.align='center'}
library(UScensus2010)
library(UScensus2010cdp)
library(ggmap)

## Build city boundary
mpls<-city("minneapolis","MN")
mpls_boundary<-fortify(mpls)

## Plot MPLS and Boundary
qmap(location = "Minneapolis, MN", zoom = 10)+
    geom_polygon(data = mpls_boundary, aes(x = long, y = lat),color = "red", alpha=0, size =1)+
    coord_map()
```

### Census Blocks
```{r, cache=FALSE,fig.width = 4.5, fig.height = 4.5,fig.align='center'}
load("data/mpls_blk.rda")
mpls_blk<-fortify(mpls_blk)

ggplot()+
   geom_polygon(data = mpls_blk, aes(x = long, y = lat, group=group),color = "blue", alpha=.2, size =.25)+
    coord_map()
```

### Neighborhoods
```{r, cache=FALSE,fig.width = 8, fig.height = 8,fig.align='center'}
load("data/mpls_neighborhoods.rda")
mpls_nb<-fortify(mpls_neighborhoods)
mpls_nb_centroid<-data.frame(coordinates(mpls_neighborhoods),name=mpls_neighborhoods$BDNAME,stringsAsFactors =FALSE)


ggplot()+
   geom_polygon(data = mpls_nb, aes(x = long, y = lat, group=group),color = "blue", alpha=.2, size =.25)+
   geom_text(aes(x = X1, y = X2,label = name),size=1.5,color="black",data=mpls_nb_centroid)+
   coord_map()
```

### School Districts

```{r, cache=FALSE,fig.width = 6, fig.height = 6,fig.align='center'}
load("data/school_attendance_dists.rda")
mpls<-spTransform(mpls,proj4string(school_attendance_dist))
mpls_sad<-gIntersection(school_attendance_dist,mpls,byid = TRUE,drop_lower_td=TRUE)
mpls_sad<-as(mpls_sad,"SpatialPolygonsDataFrame")
data<-over(mpls_sad,school_attendance_dist)
mpls_sad@data<-data


mpls_sad_centroid<-data.frame(coordinates(mpls_sad),name=tolower(mpls_sad$HIGH_NAME),stringsAsFactors =FALSE)
mpls_sad<-fortify(mpls_sad)

ggplot()+
   geom_polygon(data = mpls_sad, aes(x = long, y = lat, group=group),color = "blue", alpha=.2, size =.25)+
   geom_text(aes(x = X1, y = X2,label = name),size=2,color="black",data=mpls_sad_centroid)+
   coord_map()+
  ggtitle("High School Districs Minneapolis, MN")
```

### Crime

```{r, cache=FALSE,fig.width = 6, fig.height = 6,fig.align='center'}
load("data/crime_10_13.rda")

mpls_crime<-data.frame(coordinates(crime_10_13),stringsAsFactors =FALSE)
umn<-qmap(location = "Minneapolis, MN", zoom = 14)

umn+
   geom_point(data = mpls_crime, aes(x = coords.x1, y = coords.x2),color = "red", alpha=.5, size =.25)+
   coord_map()+
  ggtitle("Crime Locations from 2010-2013, Minneapolis MN")
```


### Police Precincts 

```{r, cache=FALSE,fig.width = 6, fig.height = 6,fig.align='center'}
load("data/mpls_PolicePrecincts.rda")

mpls_pp<-fortify(mpls_PolicePrecincts)
umn<-qmap(location = "Minneapolis, MN", zoom = 10)

umn+
   geom_polygon(data = mpls_pp, aes(x = long, y = lat, group=group),color = "red", alpha=.5, size =.25)+
   coord_map()+
  ggtitle("Police Precincts, MN")
```
