---
title: "Spatial Analysis and Mapping Workshop"
author: "Zack W Almquist (University of Minnesota)"
date: "November 8, 2016"
output: pdf_document
---

# Preliminaries
```{r,include=FALSE}
## Useful http://www.stat.cmu.edu/~cshalizi/rmarkdown/
require(knitr)
# Set knitr options for knitting code into the report:
# - Don't print out code (echo)
# - Save results so that code blocks aren't re-run unless code changes (cache),
# _or_ a relevant earlier code block changed (autodep), but don't re-run if the
# only thing that changed was the comments (cache.comments)
# - Don't clutter R output with messages or warnings (message, warning)
  # This _will_ leave error messages showing up in the knitted report
opts_chunk$set(echo=TRUE,
               cache=TRUE, autodep=TRUE, cache.comments=FALSE,
               message=FALSE, warning=FALSE,tidy=TRUE)
options(width=60)
```


This workshop will involve a large collection of R packages and will provide a basic run through for accessing spatial data, spatial plotting and integrating with the 'ggplot' packages. Last we will perform some simple spatial statistics.

## From CRAN

### Required

```{r, eval=FALSE}
install.packages("sp")
install.packages("rgeos")
install.packages("maptools")
install.packages("spdep")
install.packages("ggplot2")
install.packages("devtools")
install.packages("plyr")
```

### Recommended
```{r, eval=FALSE}
install.packages("rgdal")
```

## US Census Packages
In this workshop we will use a series of different packages, including the US Census suite of software. The most recent versions are available <https://r-forge.r-project.org/R/?group_id=2022>. We will need only three of the packages:

```{r, eval=FALSE}
install.packages("UScensus2010", repos="http://R-Forge.R-project.org")
install.packages("UScensus2010county", repos="http://R-Forge.R-project.org")
install.packages("UScensus2010tract", repos="http://R-Forge.R-project.org")
install.packages("UScensus2010cdp", repos="http://R-Forge.R-project.org")
```


### Load the US Census packages
```{r}
library(UScensus2010)
library(UScensus2010county)
data(washington.county10)
```

### Plot Washington Counties

```{r, fig.width = 4.5, fig.height = 4.5,fig.align='center'} 
plot(washington.county10)
```

## Zillow Neighborhoods

We are going to use some data from <http://zillow.com> which will allow us to work with the neighborhoods of Seattle! You can download the neighborhoods directly at <http://www.zillow.com/howto/api/neighborhood-boundaries.htm>.

### Package requirements
This function requires 'rgdal', if you were not able to instal 'rgdal' the data file is provided on the github repository as 'seattle_nb.rda'.

```{r}
require(rgdal)
```

### Function for downloading Zillow Neighborhoods
```{r}
readZillow_NB<-function(address,state,city=NULL){
  fileName<-paste(address,"ZillowNeighborhoods-",state,".zip",sep="") ##full address of zip file
  zipFile<-paste("ZillowNeighborhoods-",state,".zip",sep="")
  zipdir <- tempfile() ### Create temp file
  dir.create(zipdir) ### Create a folder in the temp file
  download.file(fileName,destfile=paste(zipdir,zipFile,sep="/"))
  unzip(paste(zipdir,zipFile,sep="/"), exdir=zipdir)
  files <- list.files(zipdir)
  sp_temp<-rgdal::readOGR(dsn=zipdir,layer=strsplit(files[grep("shp",paste(files,"0",sep=""))],"\\.")[[1]][1])
  if(!is.null(city)){
  sp_temp_city<-sp_temp[sp_temp$CITY==city,]
  return(sp_temp_city)
  }
  sp_temp
}
```
    
### Zillow Logo

Zillow requires that you add their logo to any plot using their data. Below is a quick script for doing such a task. This requires the package 'caTools'.

```{r, eval=FALSE}
install.packages('caTools')
install.packages('RCurl')
```

```{r,fig.width = 3, fig.height = 3,fig.align='center'}
library(caTools)

zillowurl<-"http://www.zillowstatic.com/vstatic/70a941d/static/logos/Zillow_Logo_HoodsProvided_RightAligned.gif"
zillow_logo <-read.gif(zillowurl,flip=TRUE)

## Plot the logo!
image(zillow_logo$image, col=zillow_logo$col, breaks=(0:length(zillow_logo$col))-0.5, asp=.15,useRaster=TRUE,axes = FALSE)
```


### Seattle Neighborhoods
```{r}  
base_address<-"http://www.zillow.com/static/shp/"
state<-"WA"
city<-"Seattle"
seattle_nb<-readZillow_NB(base_address,state,city)
```

### Plot Seattle Neighborhoods

```{r, fig.width = 4.5, fig.height = 4.5,fig.align='center'} 
par(mfrow=c(1,1),mar=c(0, 0, 0, 0) + 0.1)
plot(seattle_nb)
par(fig=c(.25, 3.3, .4, 2.5)/10,new=TRUE)
image(zillow_logo$image, col=zillow_logo$col, breaks=(0:length(zillow_logo$col))-0.5, asp=.15,useRaster=TRUE,axes = FALSE)
```

# Plot the US using ggplot

```{r,cache=FALSE} 
library(UScensus2010)
library(ggplot2)
library(dplyr)
library(plyr)
data(states)
cont_states<-states[!(states$acronym%in%c("HI","AK")),]
cont_states@data$id <- rownames(cont_states@data)
states.points = fortify(cont_states, region="id")
states.df = join(states.points, cont_states@data, by="id")
```

```{r, fig.width = 4.5, fig.height = 4.5,fig.align='center'} 
ggplot() +
  geom_polygon(data = states.df, 
               aes(x = long, y = lat, group = group, fill = pop1970-pop1960), 
               color = "black", size = 0.25) + 
  coord_map()
```


